Standardized crypto-loans on\\ the Cardano blockchain

### 2021-02-04 papers/actus2021/article.tex
Reviewed Section 5.

\textbf{TO DO:}
\begin{enumerate}
\item
Deal with inline comments in source code: \verb+% SJT ...+
\item
Write introduction [SJT]
\item
Should we include Alex as co-author?
\item
What about R (shiny)?
\end{enumerate}

@@

\subsection{Contract terms representation and explicit applicability}

@@

\section{Generating Marlowe contracts from ACTUS specs}%standardised ACTUS contract terms}
\label{generation}

@@

Generated contract is essentially a continuation chain of smaller
contracts:

\noindent 
\begin{equation}
chainlink(t)=receiveData(t)\circ \mbox{\emph{calculatePayoff(t)}}\circ \mbox{\emph{processPayoff(t)}}
\end{equation}

\noindent 
\begin{equation}
contract(ct)=collaterals(ct)\circ INIT(ct)\circ\prod_{t\in SCHED(ct)}chainlink(t)
\end{equation}

where $receiveData$ - is asking an oracle for Marlowe Choice if needed,
$\mbox{\emph{calculatePayoff}}$ - calculates payoff formula. For fixed-rate contracts
this is optimised into a pre-calculated constant. $\mbox{\emph{processPayoff}}$
- awaits Deposit of payoff amount from a party: if deposit is transferred
it redirects the funds to a counterparty - otherwise it transfers
the collateral to a counterparty and closes the contract. The chain
is generated from the fixed (known in advance) schedule of events
using the $SCHED()$ function from ACTUS.

@@

Marlowe doesn't allow functions and recursion. Naive usage of If operator
in Marlowe could lead to exponential growth of a contract, for example:

\begin{verbatim}
if condition 
	then 
		perform_something1()
		continue() 
	else 
		perform_something2 
		continue()
\end{verbatim}
\noindent
would inline contents of \texttt{continue()} twice and given that ACTUS
contracts are essentially generated using continuation as an accumulator
- this would lead to exponential explosion of the size of any ACTUS
contract that has conditionals in their state transition logic.

Example of such logic would be cap/floor limitations on interest rate:

\noindent 
\begin{equation}
adjusted=max(min(original,floor),cap)
\end{equation}

We addressed this issue by introducing the \texttt{Cond} construct in order
to represent conditional expressions. So instead of using the \texttt{If}
contract to decide the value of some variable - we use conditional
assignment instead as \texttt{Cond} is essentially just a pure function
that returns value depending on condition in contrast to the \texttt{If}
contract that performs action instead of returning value. The name
is borrowed from a similar function in Lisp.

@@

Marlowe doesn't allow perpetual contracts even if their recursion
is productive (example: perpetual swap) - thus we can't support certain
contract types from ACTUS specification, namely the ones that don't
have a defined maturity date (like UMP). In a long-term perspective
- there is a possible workaround: contracts with no maturity date
could be represented as actors with finite amounts of state transitions.
We prototyped this approach in a separate branch, however it does
seem more prone to errors comparing to rendering predefined schedules
and more importantly it greatly affects static analysis as the amount
of ``reduction steps'' grows from 
\noindent 
\begin{equation}
N_{scheduled}=count(event)
\end{equation}
to 
\noindent 
\begin{equation}
N_{stateTransitions}=\frac{max(date(event))-min(date(event))}{precision} 
\end{equation}

@@

Marlowe only supports Integers, while ACTUS is expressed in terms
of real numbers. In order to model a real number in such a setup -
we rely on fixed point precision. The algebra looks like this: 
\begin{verbatim}
(+)   = AddValue
(-)   = SubValue
a * b = Scale (1 % marloweFixedPoint) $ MulValue a b
\end{verbatim}

@@

Unfortunately, Marlowe DSL doesn't support any notion of records,
the variables can only be of type ``Integer''. Thus, in order to
map contract state (ContractStatePoly) we first pack a set of Marlowe
variables (Value Observation) representing previous $t-1$ state into
ContractStatePoly, then we apply polymorphic state transition, and
finally unpack ContractState poly into a set of Marlowe variables
into $t+1$ state:

\noindent 
\begin{equation}
st_{t+1}=unpack(stateTransition(pack(st_{t-1}))),
\end{equation}

,where pack is a chain of ``UseValue'' and unpack is a chain of
``Let''

\paragraph{Representing state transitions in Marlowe (unfolding)}

Marlowe doesn't support mutable variables, so we have to represent
state at t ($st_{t}$) literally through naming convention:

@@

\begin{figure} 
\includegraphics[width=0.8\textwidth]{images/labs.png}
\caption{Actus Labs - online tool for generating Actus contracts for Marlowe.} 
\label{fig3} 
\end{figure}

In order to demonstrate and test the capabilities of Actus generators,
a visual online Blockly-based tool was developed for Marlowe Playground.
The Actus Labs tool allows users to construct contract terms visually,
generate a corresponding Marlowe contract and try it out in a simulated
environment.

>>>>

\textbf{TO DO:}
\begin{enumerate}
\item
Deal with inline comments in source code: \verb+% SJT ...+
\item
Write introduction [SJT]
\item
Should we include Alex as co-author?
\item
What about R (shiny)?
\item
We use \verb+equation+ environment which numbers equations, but the numbering isn't used I think.
So I'd suggest moving to \verb+equation*+ instead (easy edit).
\end{enumerate}

@@

\subsection{Contract term representation and explicit applicability}

@@

\section{Generating Marlowe contracts from standardised ACTUS contract terms}
\label{generation}

In this section we describe how concrete Marlowe contracts are generate from the terms -- i.e. parameters -- of standard ACTUS contracts; we also describe the implementation of the system, and reflect on the limitations of generating contracts in Marlowe, where contracts have a a predefined lifetime and collection of interactions with contract participants.

@@

A generated contract is essentially a continuation chain of smaller contracts:

\noindent 
\begin{equation}
\ident{chainlink}(t)=\ident{receiveData}(t)\circ \ident{calculatePayoff(t)}\circ \ident{processPayoff(t)}
\end{equation}

\noindent 
\begin{equation}
\ident{contract}(ct)=\ident{collaterals}(ct)\circ \ident{INIT}(ct)\circ\prod_{\footnotesize t\in \ident{SCHED(ct)}}\ident{chainlink(t)}
\end{equation}

\noindent
where the component \emph{receiveData} asks an oracle for Marlowe Choice if needed, and
\mbox{\emph{calculatePayoff}} calculates the payoff formula. For fixed-rate contracts
this is optimised into a pre-calculated constant function. The \emph{processPayoff} function
awaits the \texttt{Deposit} of a payoff amount from a party: if the deposit is made them 
it directs the funds to a counterparty, otherwise it transfers
the collateral to the counterparty and closes the contract. The principal components of the system are shown in Figure \ref{fig1}.
% SJT added this figure ref: OK?
% Do we need to add a little more to explain Fig1?

The chain
is generated from the fixed schedule of events, known in advance of execution, 
using the \ident{SCHED()} function from ACTUS, as shown in Figure \ref{fig2}.
% SJT added this figure ref: OK?

@@

Marlowe contracts are finite, and in particular Marlowe itself does not have constructs for functions or recursion; these \emph{are} available in the Haskell and JavaScript embeddings of Marlowe, but they are unrolled on translation to pure Marlowe. Naive usage of the \texttt{If} operator in Marlowe could lead to exponential growth of a contract, as in this pesudocode example:

\begin{verbatim}
if condition 
	then 
		perform\_something1()
		continue() 
	else 
		perform\_something2() 
		continue()
\end{verbatim}
\noindent
Translating this to Marlowe would inline contents of \texttt{continue()} twice and, given that ACTUS
contracts are essentially generated using continuation as an accumulator,
this would lead to exponential explosion of the size of any ACTUS
contract that has conditionals in their state transition logic.

An example of such logic would be cap/floor limitations on interest rates:

\noindent 
\begin{equation}
\ident{adjusted}=\ident{max(min(original,floor),cap)}
\end{equation}
\noindent
We addressed this issue by introducing the \texttt{Cond} expression construct in order
to represent conditional expressions, rather than only conditional contracts as was the case before. Instead of using the \texttt{If}
contract to decide the value of some variable, we use a conditional expression instead.  
\texttt{Cond} is  a pure function
that returns a value depending on a condition, in contrast to the \texttt{If}
contract that chooses between two continuation contracts. 
% SJT I don't think we need to say this.
%The name is borrowed from a similar function in Lisp.

@@

Marlowe doesn't allow contracts that run indefinitely, even if their recursion
is productive, as would be the case in a perpetual swap contract, for example. We therefore cannot support certain
contract types from ACTUS specification, namely the ones that don't
have a defined maturity date (like UMP).

% In a long-term perspective
There is a possible workaround: contracts with no maturity date
could be represented as actors with a finite number of state transitions.
% SJT is this ^^^ rephrasing OK?
We prototyped this approach, % in a separate branch, 
however it does
seem more prone to errors comparing to rendering predefined schedules.
More importantly, it greatly affects static analysis because the number
of reduction steps in the contract grows from 
\noindent 
\begin{equation*}
N_{\footnotesize \ident{scheduled}}=\ident{count(event)}
\end{equation*}
to 
\noindent 
\begin{equation}
N_{\footnotesize \ident{stateTransitions}}=\frac{\ident{max(date(event))}-\ident{min(date(event))}}{\ident{precision}} 
\end{equation}

@@

For numeric types Marlowe supports Integers, while ACTUS is expressed in terms
of real numbers. In order to model a real number in such a setup
we rely on fixed-point precision. The algebra looks like this: 
\begin{verbatim}
  (+)   = AddValue
  (-)   = SubValue
  a * b = Scale (1 % marloweFixedPoint) $ MulValue a b
\end{verbatim}
% SJT this needs more explanation here ^^^
\noindent
We plan to move Marlowe to fixed-precision numbers for the on-blockchain implementation available later in 2021.

@@

The Marlowe DSL does not support any notion of records,
the variables can only be of type ``Integer''. In order to
map contract state -- \emph{ContractStatePoly} -- we pack a set of Marlowe
variables of type \texttt{Value} and \texttt{Observation}, representing the previous ($t-1$) state, passing them into
\emph{ContractStatePoly}, to apply the polymorphic state transition, and
finally unpack \emph{ContractStatePoly} into a set of Marlowe variables representing the 
 state at $t+1$:
% SJT again I am confused about t-1 to t+1 transition. Why not just one step?  
\noindent 
\begin{equation}
st_{t+1}=unpack(stateTransition(pack(st_{t-1}))),
\end{equation}
\noindent
where \emph{pack} is a chain of \texttt{UseValue} constructs and \ident{unpack} is a chain of
\texttt{Let}s.

\paragraph{Representing state transitions in Marlowe}

Marlowe is a declarative language, and so in partticular does support mutable variables. We therefore represent the state at stage $t$ ($st_{t}$) literally through this naming convention:

@@

\begin{figure}[t]
\includegraphics[width=0.8\textwidth]{images/labs.png}
\caption{Actus Labs - an online tool for generating Actus contracts for Marlowe.} 
\label{fig3} 
\end{figure}

In order to demonstrate and test the capabilities of Actus generators,
a visual online Blockly-based tool was developed for the Marlowe Playground.
The Actus Labs tool, shown in Figure \ref{fig3}, allows users to construct contract terms visually
to generate a corresponding Marlowe contract and then to try it out in a simulation 
environment.
